Exploring the Pareto front of multi-objective COVID-19 mitigation policies using reinforcement learning

### 2023-03-31 paper/main.tex
pieter comments

\newcommand{\foi}{\lambda}
\newcommand\todo[1]{\textcolor{red}{TODO: #1}}

%MDP
%state space
\newcommand{\mdpstatespace}{\mathcal{S}}
%action space
\newcommand{\mdpactionspace}{\mathcal{A}}
%mdp discount factor
\newcommand{\mdpdiscount}{\gamma}
%mdp policy
\newcommand{\mdppolicy}{\pi}
%mdpvaluefunction
\newcommand{\mdpvaluefunction}{V}
%vectorial value function
\newcommand{\momdpvaluefunction}{\mathbf{V}}
%mdpqfunction
\newcommand{\mdpqfunction}{Q}
%mdp q function estimate
\newcommand{\mdpqfunctionestimate}{\hat{Q}}
%mdp optimal q function
\newcommand{\mdpoptimalqfunction}{Q^{*}}
%q step size
\newcommand{\qstepsize}{\alpha}
%\advantageestimate
\newcommand{\advantageestimate}{\hat{A}}
%transition function
\newcommand{\mdptransition}{\mathcal{T}}
%reward fn
\newcommand{\mdprewardfn}{\mathcal{R}}
%vectorial reward fn
\newcommand{\momdprewardfn}{\bm{\mathcal{R}}}
%mdp state
\newcommand{\state}{s}
\newcommand{\mdpstate}{\mathbf{s}}
%mdp reward
\newcommand{\mdpreward}{r}
\newcommand{\momdpreward}{\mathbf{r}}
\newcommand{\momdpreturn}{\mathbf{R}}
%mdp action
\newcommand{\action}{a}
%vectorial action
\newcommand{\mdpaction}{\mathbf{a}}
% the 6 social contact matrices
\newcommand{\ctotal}{C}
\newcommand{\chome}{C_{\text{home}}}
\newcommand{\cwork}{C_{\text{work}}}
\newcommand{\ctransport}{C_{\text{transport}}}
\newcommand{\cschool}{C_{\text{school}}}
\newcommand{\cleisure}{C_{\text{leisure}}}
\newcommand{\cother}{C_{\text{other}}}
% tentative momdp env name
\newcommand{\momdpname}{MOBelCov}
%other params
\newcommand{\proportionalityfactor}{q}
\newcommand{\agegroup}{k}
\newcommand{\agegroups}{K}
\newcommand{\budget}{\bm{b}}

@@

\author{Mathieu Reymond \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\texttt{mathieu.reymond@vub.be} \\
	\And
	Conor F. Hayes \\
	National University of Ireland Galway\\
	Galway, Ireland\\
% 	\texttt{c.hayes13@nuigalway.ie} \\
	\And
	Lander Willem \\
	University of Antwerp \\
	Antwerp, Belgium \\
	\And
	Roxana R\u{a}dulescu \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Steven Abrams \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Diederik M.\ Roijers \\
	HU University of Applied Sciences Utrecht \\
	Utrecht, the Netherlands \\
	\And
	Enda Howley \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Patrick Mannion \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Niel Hens \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Ann Now\'{e} \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Pieter Libin \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
% 	\AND
% 	Coauthor \\
% 	Affiliation \\
% 	Address \\
% 	\texttt{email} \\
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Infectious disease outbreaks represent a major global challenge \todo{ref}. To this end, understanding the complex dynamics that underlie these epidemics is essential. Epidemiological transmission models allow us to capture and understand such dynamics and facilitate the study of prevention strategies through simulation. However, developing efficient mitigation strategies remains a challenging process, given the non-linear and complex nature of epidemics.

To address these challenges, reinforcement learning provides a methodology to automatically learn mitigation strategies in combination with complex epidemic models~\cite{libin2020}.
Previous research focused on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost). Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
Typically, a decision maker will rely on an expert's assistance to manually design such a metric. However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits sub-optimal performance. Furthermore, manually designing such metrics is time consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}. Moreover, taking a single objective approach has several other limiting factors. For example, it reduces the explainability of the learned solution, as we cannot compare the learned behavior with alternatives. Moreover, the preferences of the decision maker might change over time, in which case the optimization process needs to be done again~\cite{hayes2021practical}.

This challenging process can be circumvented by taking an explicit multi-objective approach that aims to learn the different trade-offs regarding the considered criteria. By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}. This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that each solution implies.

In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective epidemic mitigation strategies. We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a strict lockdown~\cite{willem2021impact}. When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe contact and mobility reductions.
Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic, and was fitted to hospitalisation incidence data and serial sero-prevalence data~\citet{abrams2021modelling}.
This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.  
Based on this model, we contribute a novel mutli-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}. \momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.

To learn and explore the trade-offs between the attack rate and social burden we present a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}. PCN uses a single neural network to learn the policies that belong to the Pareto front.
As PCN is an algorithm designed for discrete action-spaces, we extend it towards continuous action-spaces to accommodate \momdpname's action-space. With this continuous action variant of PCN, we explore the Pareto front of multi-objective COVID-19 mitigation policies.

By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low \todo{update (challenges)}. Therefore, in this work we illustrate that multi-objective reinforcement learning can be utilised to provide important insights surrounding the trade-offs between complex mitigation polices in real-world epidemiological models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
In the following, we use bold notation $\momdpvaluefunction$ to denote a vectorial value, while non-bold notation $\mdpvaluefunction$ for its scalar counterpart.

@@

While MOMDPs define the actions an agent can take, it does not define the agent's behavior, i.e., which actions are taken in each state. Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\state \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$. We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\state_0$ until the end of the decision problem:

\begin{equation}
    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \state_0 \right ],
\end{equation}
%
where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\state_t, \action_t, \state_{t+1})$, where $\state_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\state_{t+1}$ signifies the state reached by taking action $\action_t$.

For single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
\begin{equation}
\mdppolicy^*= \arg\max_\mdppolicy \mdpvaluefunction^\mdppolicy.
\end{equation}

In contrast, for MORL, $n > 1$ which leads to vectorial returns. In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other. For example, it is impossible to decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if both policies lead to expected returns $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively. We call these solutions \emph{non-dominated}, i.e., solutions for which it is impossible to improve an objective without hampering another. The set that contains all the non-dominated solutions of the decision problem is called the \emph{Pareto front} $\mathcal{F}$. Our goal is to find the set of policies that lead to all the $\mdpvaluefunction$-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$. In general, we call any set of $\mdpvaluefunction$-values a \emph{solution set}. When a solution set contains only  non-dominated $\mdpvaluefunction$-values, it is referred to as a \emph{coverage set}. In the case that no $\mdppolicy$ exists that has a $\momdpvaluefunction^{\pi}$ dominating any of the solutions in a coverage set, then this coverage set is the Pareto front.

\subsection{Multi-Objective Metrics}
Comparing the learned coverage sets produced by different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another. Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences. In this work we use several metrics to evaluate our algorithm's performance.

@@

\begin{equation}
\label{eq:epsilon-metric}
    I_{\varepsilon} = 
    \inf_{\varepsilon\in\mathbb{R}} 
    \{ 
    \forall\ \momdpvaluefunction^\pi\!{\in}\ \mathcal{F}, ~
    \exists\ \momdpvaluefunction^{\pi'}\!{\in}\ \hat{\Pi} :\ 
    || \momdpvaluefunction^\pi - \momdpvaluefunction^{\pi'} ||_{\infty} \le \varepsilon
    \},
\end{equation}
%
where $||.||_{\infty}$, the L-infinity norm, returns the magnitude of the largest entry of a vector.

@@

We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., it will still report low performance as long as a single point of the Pareto front is incorrectly modeled, even if all the other points are covered. As such, we also use the $I_{\varepsilon-mean}$ \cite{reymond2022pcn} which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set.

@@

\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
      \begin{axis}[legend pos=south west, xmin=-1, xmax=4.5, ymin=-0.5, ymax=4.5,ytick={0,1,2,3,4,5},scale only axis,width=0.4\textwidth]
      \addplot[only marks,mark=x,mark size=3] coordinates {(-0.5,0)};
      \addplot[only marks,mark=*] coordinates {(1,4)(2,2)(4,1)};
      \addplot[only marks,mark=o] coordinates {(0.5,3)(0.75,2.3)(2.3,1)(3.3,0.7)};
      % Hypervolume
      \fill [cyan,opacity=0.2] (-0.5,0) -- (-0.5,3) -- (0.5,3) -- (0.5,2.3) -- (0.75,2.3) --
      (0.75,1) -- (2.3,1) -- (2.3,0.7) -- (3.3,0.7) -- (3.3,0) -- cycle;
      % epsilons
      \draw [dashed, -|] (0.5, 3) -- node[right]{$\varepsilon_1$} (0.5,4);
      %\draw [dashed, -|] (0.75,2.3) -- node[below]{x} (2,2.3);
      \draw [dashed, -|] (2.3,1) -- node[right]{$\varepsilon_2$} (2.3,2);
      \draw [dashed, -|] (3.3,0.7) -- node[below]{$\varepsilon_3$} (4,0.7);
      \draw [opacity=0.2,line width=2,orange, dotted] (-0.5,0) -- (-0.5,4) -- (1,4) -- (1,2) --
      (2,2) -- (2,1) -- (4,1) -- (4,0) -- (-0.5,0);
      \end{axis}
    \end{tikzpicture}
    \caption{Example of a Pareto front (black dots) and a coverage set (white dots) in a 2-objective environment. The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to some reference point (cross). The reference point is taken as a lower bound on the achievable returns, which is why it can differ from the origin. The $\varepsilon$ metrics first compute the maximum distance between each point in the Pareto front and its closest point in the coverage set ($\varepsilon_i$). We can then take their maximum value to compute the $I_\varepsilon$ metric, or their mean value to obtain the $I_{\varepsilon-mean}$ metric of the coverage set.}
    \label{fig:paretofrontmetrics}
\end{figure}

@@

A visual representation of the model is depicted in Fig~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move from one compartment to another.

@@

\begin{align*}
\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
%
\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
%
\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
%
\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
%
\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
%
\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
%
\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
%
\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
\end{align*}
%
where, for example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$.

In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$. Infection dynamics are governed by an age-specific force of infection $\foi$:

\begin{equation}
    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
\end{equation}
%
where $\agegroup$ is the respective age group of a total of $\agegroups$ age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.

@@

Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$. Therefore, it is possible to define the transmission rates for both symptomatic and asymptomatic individuals as follows:

@@

Moreover, the age-dependent force of infection can be defined as follows:

>>>>

\newcommand{\foi}{\lambda}
\newcommand\todo[1]{\textcolor{red}{(TODO: #1)}}

%MDP
%state space
\newcommand{\mdpstatespace}{\mathcal{S}}
%action space
\newcommand{\mdpactionspace}{\mathcal{A}}
%mdp discount factor
\newcommand{\mdpdiscount}{\gamma}
%mdp policy
\newcommand{\mdppolicy}{\pi}
%mdpvaluefunction
\newcommand{\mdpvaluefunction}{V}
%vectorial value function
\newcommand{\momdpvaluefunction}{\mathbf{V}}
%mdpqfunction
\newcommand{\mdpqfunction}{Q}
%mdp q function estimate
\newcommand{\mdpqfunctionestimate}{\hat{Q}}
%mdp optimal q function
\newcommand{\mdpoptimalqfunction}{Q^{*}}
%q step size
\newcommand{\qstepsize}{\alpha}
%\advantageestimate
\newcommand{\advantageestimate}{\hat{A}}
%transition function
\newcommand{\mdptransition}{\mathcal{T}}
%reward fn
\newcommand{\mdprewardfn}{\mathcal{R}}
%vectorial reward fn
\newcommand{\momdprewardfn}{\bm{\mathcal{R}}}
%mdp state
\newcommand{\mdpstate}{\mathbf{s}}
%mdp reward
\newcommand{\mdpreward}{r}
\newcommand{\momdpreward}{\mathbf{r}}
\newcommand{\momdpreturn}{\mathbf{R}}
%mdp action
\newcommand{\action}{a}
%vectorial action
\newcommand{\mdpaction}{\mathbf{a}}
% the 6 social contact matrices
\newcommand{\ctotal}{C}
\newcommand{\chome}{C_{\text{home}}}
\newcommand{\cwork}{C_{\text{work}}}
\newcommand{\ctransport}{C_{\text{transport}}}
\newcommand{\cschool}{C_{\text{school}}}
\newcommand{\cleisure}{C_{\text{leisure}}}
\newcommand{\cother}{C_{\text{other}}}
% tentative momdp env name
\newcommand{\momdpname}{MOBelCov}
%other params
\newcommand{\proportionalityfactor}{q}
\newcommand{\agegroup}{k}
\newcommand{\agegroups}{K}
\newcommand{\budget}{\bm{b}}

@@

\author{Mathieu Reymond \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\texttt{mathieu.reymond@vub.be} \\
	\And
	Conor F. Hayes \\
	National University of Ireland Galway\\
	Galway, Ireland\\
% 	\texttt{c.hayes13@nuigalway.ie} \\
	\And
	Lander Willem \\
	University of Antwerp \\
	Antwerp, Belgium \\
	\And
	Roxana R\u{a}dulescu \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Steven Abrams \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Diederik M.\ Roijers \\
	HU University of Applied Sciences Utrecht \\
	Utrecht, the Netherlands \\
	\And
	Enda Howley \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Patrick Mannion \\
	National University of Ireland Galway\\
	Galway, Ireland\\
	\And
	Niel Hens \\
	Hasselt University \\
	Hasselt, Belgium \\
	\And
	Ann Now\'{e} \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
	\And
	Pieter Libin \\
	Vrije Universiteit Brussel\\
	Brussels, Belgium \\
    \texttt{pieter.libin@vub.be} \\
% 	\AND
% 	Coauthor \\
% 	Affiliation \\
% 	Address \\
% 	\texttt{email} \\
}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Infectious disease outbreaks represent a major global challenge \cite{miranda2022tale}. To this end, understanding the complex dynamics that underlie these epidemics is essential. Epidemiological transmission models allow us to capture and understand such dynamics and facilitate the study of prevention strategies through simulation. However, developing efficient mitigation strategies remains a challenging process, given the non-linear and complex nature of epidemics.

To address these challenges, reinforcement learning provides a methodology to automatically learn mitigation strategies in combination with complex epidemic models~\cite{libin2020}.
Previous research focused on optimising policies with respect to a single objective, such as the pathogen's attack rate, while the mitigation of epidemics is a problem that inherently covers distinct  and possibly conflicting criteria (i.a., prevalence, mortality, morbidity, cost). Therefore, optimizing on a single objective requires that  these distinct criteria are somehow aggregated into a single metric.
Typically, a decision maker will rely on an expert's assistance to manually design such a metric. However, this is a challenging process, as the decision maker is generally not aware of all possible trade-offs, and as such this may result in a metric that exhibits suboptimal performance. Furthermore, manually designing such metrics is time-consuming, costly and error-prone, as this non-intuitive process requires repetitive and tedious tuning to achieve the desired behaviour~\cite{roijers2013survey}. Moreover, taking a single objective approach has several other limiting factors. For example, it reduces the explainability of the learned solution, as we cannot compare the learned behavior with alternatives. Moreover, the preferences of the decision maker might change over time, in which case the optimization process needs to be done again~\cite{hayes2021practical}.

This challenging process can be circumvented by taking an explicit multi-objective approach that aims to learn the different trade-offs regarding the considered criteria. By assuming that a decision maker will always prefer solutions for which at least one objective improves, it is possible to learn a set of optimal solutions referred to as the \emph{Pareto front}~\cite{hayes2021practical}. This enables decision makers to review each solution on the Pareto front before making a decision, thereby being aware of the trade-offs that each solution implies.

In this work, we investigate the use of \emph{multi-objective reinforcement learning} (MORL) to learn a set of solutions that approximate the Pareto front of multi-objective epidemic mitigation strategies. We consider the first wave of the Belgian COVID-19 epidemic, which was mitigated by a strict lockdown~\cite{willem2021impact}. When the incidence of confirmed cases was steadily decreasing, epidemiological experts were tasked to investigate deconfinement strategies, to reduce the severe social contact and mobility restrictions.
Here, we consider an epidemiological model that was constructed to describe the Belgian COVID-19 epidemic and was fitted to hospitalisation incidence data and serial sero-prevalence data~\cite{abrams2021modelling}.
This model constitutes a stochastic discrete-time age-structured compartmental model that simulates mitigation strategies by varying social distancing parameters concerning school, work and leisure contacts.  
Based on this model, we contribute a novel mutli-objective epidemiological reinforcement learning environment~(\momdpname), in the form of a multi-objective Markov decision process (MOMDP) \cite{roijers2013survey}. \momdpname\ encapsulates the epidemiological model developed by \citet{abrams2021modelling} to implement state transitions, with an action space that combines a proportional reduction of school, work and leisure contacts at each time step, Furthermore, it defines a reward function based on two objectives: the attack rate (i.e., proportion of the population affected by the pathogen) and the social burden that is induced by the mitigation measures.

To learn and explore the trade-offs between the attack rate and social burden we use a state-of-the-art MORL approach based on Pareto Conditioned Networks (PCN)~\cite{reymond2022pcn}. PCN uses a single neural network to learn the policies that belong to the Pareto front.
As PCN is an algorithm designed for discrete action-spaces, we extend it towards continuous action-spaces to accommodate \momdpname's action-space. With this continuous action variant of PCN, we explore the Pareto front of multi-objective COVID-19 mitigation policies.

By evaluating the solution set of mitigation policies learned by PCN, we observe that PCN minimises the social burden in scenarios where hospitalization rates are sufficiently low. Therefore, in this work we illustrate that multi-objective reinforcement learning can provide important insights concerning the trade-offs between complex mitigation polices in real-world epidemiological models.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
In the following, we use bold notation to denote variables with a vectorial value, while non-bold notation for its scalar counterpart.

@@

While MOMDPs define the actions an agent can take, it does not define the agent's behavior, i.e., which actions are taken in each state. Given this MOMDP, an agent follows a policy $\mdppolicy$, that expresses the probability to take action $\action \in \mdpactionspace$ when in state $\mdpstate \in \mdpstatespace$: $\mdpstatespace \times \mdpactionspace \to \left[ 0, 1 \right]$. We measure the performance of $\mdppolicy$ through the expected sum of discounted rewards (denoted the \emph{Value}) it achieves from the initial state $\mdpstate_0$ until the end of the decision problem:

\begin{equation}
    \momdpvaluefunction^\mdppolicy = \mathbb{E} \left [ \sum_{t=0}^h \mdpdiscount^t \momdpreward_t |\ \mdppolicy, \mdpstate_0 \right ],
\end{equation}
%
where $\momdpreward_t$ corresponds to the multi-objective reward observed at time $t$, by following policy $\mdppolicy$, given by the reward function $\mdprewardfn(\mdpstate_t, \action_t, \mdpstate_{t+1})$, where $\mdpstate_t$ is the current state, $\action_t$ corresponds to the action that was taken, and $\mdpstate_{t+1}$ signifies the state reached by taking action $\action_t$.

On the one hand, for single-objective RL, where $n = 1$, the goal is to find the policy $\mdppolicy^{*}$ that maximizes the $\mdpvaluefunction$-value:
\begin{equation}
\mdppolicy^*= \arg\max_\mdppolicy \mdpvaluefunction^\mdppolicy.
\end{equation}

On the other hand, for MORL, $n > 1$ which leads to vectorial returns. In this case, there can be policies for which, without any additional information, it is impossible to know if one is better than the other. For example, it is impossible to decide which policy between $\mdppolicy_1, \mdppolicy_2$ is optimal if both policies lead to expected returns $\momdpvaluefunction^{\mdppolicy_1}=(0,1), \momdpvaluefunction^{\mdppolicy_2}=(1,0)$ respectively. We call these solutions \emph{non-dominated}, i.e., solutions for which it is impossible to improve an objective without hampering another. The set that contains all the non-dominated solutions of the decision problem is called the \emph{Pareto front} $\mathcal{F}$. Our goal is to find the set of policies that lead to all the $\mdpvaluefunction$-values contained in the Pareto front $\Pi^* = \{\ \pi | \momdpvaluefunction^{\pi} \in \mathcal{F}\}$. In general, we call any set of $\mdpvaluefunction$-values a \emph{solution set}. When a solution set contains only  non-dominated $\mdpvaluefunction$-values, it is referred to as a \emph{coverage set}. In the case that no $\mdppolicy$ exists that has a $\momdpvaluefunction^{\pi}$ dominating any of the solutions in a coverage set, then this coverage set is the Pareto front.

\subsection{Multi-Objective Metrics}
Comparing the learned coverage sets produced by different algorithms is a non-trivial task, as one algorithm's output might dominate the other in some region of the objective-space, but be dominated in another. Intuitively, one would generally prefer the algorithm that covers a wider range of decision maker preferences. %In this work we use several metrics to evaluate our algorithm's performance.

@@

\begin{equation}
\label{eq:epsilon-metric}
    I_{\varepsilon} = 
    \inf_{\varepsilon\in\mathbb{R}} 
    \{ 
    \forall\ \momdpvaluefunction^\pi\!{\in}\ \mathcal{F}, ~
    \exists\ \momdpvaluefunction^{\pi'}\!{\in}\ \hat{\Pi} :\ 
    || \momdpvaluefunction^\pi - \momdpvaluefunction^{\pi'} ||_{\infty} \le \varepsilon
    \},
\end{equation}
%
where $||.||_{\infty}$, the L-infinity norm, is defined as the magnitude of the largest entry of a vector.

@@

We note that the $\varepsilon$-indicator metric is quite pessimistic, as it measures worst-case performance~\cite{zintgraf2015}, i.e., it will still report low performance as long as a single point of the Pareto front is incorrectly modeled, even if all the other points are covered. As such, we also use the $I_{\varepsilon-mean}$ which measures the \emph{average} $\varepsilon$ distance of the solutions in $\mathcal{F}$ with respect to the evaluated coverage set~\cite{reymond2022pcn}.

@@

\begin{figure}[t!]
    \centering
    \begin{tikzpicture}
      \begin{axis}[legend pos=south west, xmin=-1, xmax=4.5, ymin=-0.5, ymax=4.5,ytick={0,1,2,3,4,5},scale only axis,width=0.4\textwidth]
      \addplot[only marks,mark=x,mark size=3] coordinates {(-0.5,0)};
      \addplot[only marks,mark=*] coordinates {(1,4)(2,2)(4,1)};
      \addplot[only marks,mark=o] coordinates {(0.5,3)(0.75,2.3)(2.3,1)(3.3,0.7)};
      % Hypervolume
      \fill [cyan,opacity=0.2] (-0.5,0) -- (-0.5,3) -- (0.5,3) -- (0.5,2.3) -- (0.75,2.3) --
      (0.75,1) -- (2.3,1) -- (2.3,0.7) -- (3.3,0.7) -- (3.3,0) -- cycle;
      % epsilons
      \draw [dashed, -|] (0.5, 3) -- node[right]{$\varepsilon_1$} (0.5,4);
      %\draw [dashed, -|] (0.75,2.3) -- node[below]{x} (2,2.3);
      \draw [dashed, -|] (2.3,1) -- node[right]{$\varepsilon_2$} (2.3,2);
      \draw [dashed, -|] (3.3,0.7) -- node[below]{$\varepsilon_3$} (4,0.7);
      \draw [opacity=0.2,line width=2,orange, dotted] (-0.5,0) -- (-0.5,4) -- (1,4) -- (1,2) --
      (2,2) -- (2,1) -- (4,1) -- (4,0) -- (-0.5,0);
      \end{axis}
    \end{tikzpicture}
    \caption{Example of a Pareto front (black dots) and a coverage set (white dots) in a 2-objective environment. The hypervolume metric (in light blue) measures the volume of all dominated solutions with respect to a reference point (cross). The reference point is taken as a lower bound on the achievable returns, which is why it can differ from the origin. The $\varepsilon$ metrics first compute the maximum distance between each point in the Pareto front and its closest point in the coverage set ($\varepsilon_i$). We can then take their maximum value to compute the $I_\varepsilon$ metric, or their mean value to obtain the $I_{\varepsilon-mean}$ metric of the coverage set.}
    \label{fig:paretofrontmetrics}
\end{figure}

@@

A visual representation of the model is depicted in Fig.~\ref{fig:compartment_model}. It shows the different compartments, as well as the flow rates at which individuals move between compartments.

@@

\begin{align*}
\frac{d\textbf{S}(t)}{dt} & = -\lambda(t)(\textbf{S})(t), \\
%
\frac{d\textbf{E}(t)}{dt} & = \lambda(t)(\textbf{S})(t) - \gamma \textbf{E}(t), \\
%
\frac{d\textbf{I}^{presym}(t)}{dt} & = \gamma \textbf{E}(t) - \theta \textbf{I}^{presym}(t), \\
%
\frac{d\textbf{I}^{asym}(t)}{dt} & = \theta p \textbf{I}^{presym}(t) - \delta_{1} \textbf{I}^{asym}(t), \\
%
\frac{d\textbf{I}^{mild}(t)}{dt} & = \theta (1 - p) \textbf{I}^{presym}(t) - \{ \psi + \omega_{2} \} \textbf{I}^{mild}(t), \\
%
\frac{d\textbf{I}^{sev}(t)}{dt} & = \psi \textbf{I}^{mild}(t) - \omega \textbf{I}^{sev}(t), \\
%
\frac{d\textbf{I}^{hosp}(t)}{dt} & = \phi_{1} \omega \textbf{I}^{sev}(t) - \{ \delta_{3} + \tau_{1} \} \textbf{I}^{hosp}(t), \\
%
\frac{d\textbf{I}^{icu}(t)}{dt} & = (1 - \phi_1) \omega \textbf{I}^{sev}(t) - \{ \delta_{4} + \tau_{2} \} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{D}(t)}{dt} & = \tau_{1} \textbf{I}^{hosp}(t) - \tau_{2} \textbf{I}^{icu}(t), \\
%
\frac{d\textbf{R}(t)}{dt} & = \delta_{1} \textbf{I}^{asym}(t) + \delta_{2} \textbf{I}^{mild}(t) + \delta_{3} \textbf{I}^{hosp}(t) + \delta_{4} \textbf{I}^{icu}(t)
\end{align*}
%

In this set of ordinary differential equations, each state variable represents a vector over all age groups for a particular compartment at time $t$. For example, $\textbf{S} = (S_{1}(t), S_{2}(t), ..., S_{\agegroup}(t))^{T}$ is the vector representing the susceptible members of the population of each age group $\agegroup$ at time $t$. Infection dynamics are governed by an age-specific force of infection $\foi$:

\begin{equation}
    \foi(\agegroup,t) = \sum_{\agegroup'=1}^{\agegroups}\beta(\agegroup,\agegroup')I_{\agegroup'}(t),
\end{equation}
%
where $\agegroups$ is the total number of age groups, and $\beta(\agegroup,\agegroup')$ is the time-invariant transmission rate that encodes the average per capita rate at which an infectious individual in age group $\agegroup$ makes an effective contact with a susceptible individual in age group $\agegroup'$, per unit of time.

@@

Following \citet{abrams2021modelling}, we rely on distinct social contact matrices for symptomatic and asymptomatic individuals, respectively $\ctotal_{s}$ and $\ctotal_{a}$. Therefore, we define the transmission rates for both symptomatic and asymptomatic individuals as follows:

@@

The age-dependent force of infection can be defined as follows:
